### 2020-09-07 old-versions/dac2020-hardwareKEM/results.tex
first commit

\section{Results} \label{sec:results}

%One of the main contributions of this research is the proposal to use an alternative PRNG to the one used by Howe et al. \cite{howe2018standard}, cSHAKE. Comparatively, 
In this section we present the results obtained when implementing our FrodoKEM architecture. We provide a table of results for each of the key generation, encapsulation, and decapsulation designs in Tables \ref{tab:keyresults}, \ref{tab:encapsresults}, and \ref{tab:decapsresults}, respectively. We also provide results for the PRNG and Gaussian sampler in Table \ref{tab:otherresults}. All tables give comparative results of the previous FrodoKEM design in hardware, which utilize 1x LWE multiplier per clock cycle and completely conform to the FrodoKEM specifications by using cSHAKE where we are using Trivium. Moreover, all results are benchmarked on the same FPGA device as previous work, Xilinx Artix-7 XC7A35T FPGA, running on Vivado 2019.1.

\begin{table}[!t]
\centering
\caption{FPGA resource consumption of the proposed FrodoKEM KeyGen designs, using $k$ parallel multipliers, for parameter sets FrodoKEM-640 and FrodoKEM-976. All results utilise a Xilinx Artix-7 FPGA.}\label{tab:keyresults}
%\vspace{-.2cm}
%\resizebox{0.7\textwidth}{!}{
\noindent\makebox[\textwidth]{
\begin{tabular}{l N{4}{0} N{4}{0} N{4}{0} N{2}{0} N{2}{0} N{3}{0} N{3}{0} }
\hline \Tstrut
  {\bf \textsc{FrodoKEM} Protocol}
&  {\bf LUT} 
&  {\bf FF} 
&  {\bf Slices}  
&  {\bf DSP} 
&  {\bf BRAM}
&  {\bf MHz} 
&  {\bf Thr-Put}  \\ \hline \Tstrut
%\textbf{} & & & \textbf{BRAM} & &\textbf{Sec} \\ \hline

KeyGen-640 1x & 971 & 433 & 290 & 1 & 0 & 191 & 59 \\
KeyGen-640 4x & 1174 & 781 & 355 & 4 & 0 & 185 & 226 \\
KeyGen-640 8x & 1679 & 1570 & 532 & 8 & 0 & 182 & 445 \\
KeyGen-640 16x & 2587 & 2994 & 855 & 16 & 0 & 172 & 840 \\ \hline \Tstrut

KeyGen-640 1x \cite{howe2018standard} & 3771 & 1800 & 1035 & 1 & 6 & 167 & 51 \\ \hline \Tstrut

KeyGen-976 1x & 1243 & 441 & 362 & 1 & 0 & 189 & 25 \\
KeyGen-976 4x & 1458 & 792 & 440 & 4 & 0 & 184 & 97 \\
KeyGen-976 8x & 1967 & 1576 & 617 & 8 & 0 & 178 & 187 \\
KeyGen-976 16x & 2869 & 3000 & 908 & 16 & 0 & 169 & 355 \\ \hline \Tstrut
KeyGen-976 1x \cite{howe2018standard} & 7139 & 1800 & 1939 & 1 & 8 & 167 & 22 \\ \hline \Tstrut

\end{tabular}}
\end{table}

\begin{table}[!t]
\centering
\caption{FPGA resource consumption of the proposed FrodoKEM Encaps designs, using $k$ parallel multipliers, for parameter sets FrodoKEM-640 and FrodoKEM-976. All results utilize a Xilinx Artix-7 FPGA.}\label{tab:encapsresults}
%\vspace{-.2cm}
%\resizebox{0.7\textwidth}{!}{
\noindent\makebox[\textwidth]{
\begin{tabular}{l N{4}{0} N{4}{0} N{4}{0} N{2}{0} N{2}{0} N{3}{0} N{3}{0} }
\hline \Tstrut
  {\bf \textsc{FrodoKEM} Protocol}
&  {\bf LUT} 
&  {\bf FF} 
&  {\bf Slices}  
&  {\bf DSP} 
&  {\bf BRAM}
&  {\bf MHz} 
&  {\bf Thr-Put}  \\ \hline \Tstrut
%\textbf{} & & & \textbf{BRAM} & &\textbf{Sec} \\ \hline

Encaps-640 1x & 4246 & 2131 & 1180 & 1 & 0 & 190 & 58 \\
Encaps-640 4x & 4620 & 2552 & 1338 & 4 & 0 & 183 & 221 \\
Encaps-640 8x & 5155 & 3356 & 1485 & 8 & 0 & 177 & 427 \\
Encaps-640 16x & 5796 & 4694 & 1692 & 16 & 0 & 171 & 825 \\ \hline \Tstrut
Encaps-640 1x \cite{howe2018standard} & 6745 & 3528 & 1855 & 1 & 11 & 167 & 51 \\ \hline \Tstrut

Encaps-976 1x & 4650 & 2118 & 1272 & 1 & 0 & 187 & 25 \\
Encaps-976 4x & 4996 & 2611 & 1455 & 4 & 0 & 180 & 94 \\
Encaps-976 8x & 5562 & 3349 & 1608 & 8 & 0 & 175 & 183 \\
Encaps-976 16x & 6188 & 4678 & 1782 & 16 & 0 & 168 & 350 \\ \hline \Tstrut

Encaps-976 1x \cite{howe2018standard} & 7209 & 3537 & 1985 & 1 & 16 & 167 & 22 \\ \hline \Tstrut

\end{tabular}}
\end{table}

@@

There is a significant increase in area consumption of all the decapsulation results which do not utilize BRAM. This is mainly due to the need of storing public-key and secret-key matrices. We provide results for both architectures with and without BRAM. The design without BRAM has a significantly higher throughput, due to the much higher frequency. These results are reported in Figure \ref{fig:hw_line}, which shows the efficiency of each design (namely their throughput) per FPGA slice utilized. Figure \ref{fig:hw} shows a slice count summary of all the proposed designs, showing a consistent and fairly linear increase in slice utilization as the number of parallel multipliers increases. We note on decapsulation results in Figure \ref{fig:hw} where the results would lie if BRAM is used, hence the total results for without BRAM include both red areas (i.e., they overlap). In most cases slice counts at least double for decapsulation when BRAM is removed, with only slight increases in throughout, hence it might be prudent in some use cases to keep BRAM usage. BRAM usage, however, is not as friendly when hardware designs are considered for ASIC, thus is it useful to consider designs both with and without BRAM.

\begin{table}[!t]
\centering
\caption{FPGA resource consumption of the proposed FrodoKEM Decaps designs, using $k$ parallel multipliers, for parameter sets FrodoKEM-640 and FrodoKEM-976. Results with * mean BRAM is used. All results utilize a Xilinx Artix-7 FPGA.}\label{tab:decapsresults}
%\vspace{-.2cm}
%\resizebox{0.7\textwidth}{!}{
\noindent\makebox[\textwidth]{
\begin{tabular}{l N{5}{0} N{4}{0} N{4}{0} N{2}{0} N{2}{1} N{3}{0} N{3}{0} }
\hline \Tstrut
  {\bf \textsc{FrodoKEM} Protocol}
&  {\bf LUT} 
&  {\bf FF} 
&  {\bf Slices}  
&  {\bf DSP} 
&  {\bf BRAM}
&  {\bf MHz} 
&  {\bf Thr-Put}  \\ \hline \Tstrut
%\textbf{} & & & \textbf{BRAM} & &\textbf{Sec} \\ \hline

Decaps-640 1x & 10518 & 2299 & 2933 & 1 & 0 & 190 & 57 \\
Decaps-640 4x & 11581 & 2818 & 3424 & 4 & 0 & 174 & 208 \\
Decaps-640 8x & 13128 & 3737 & 3710 & 8 & 0 & 164 & 391 \\
Decaps-640 16x & 14528 & 5335 & 4020 & 16 & 0 & 160 & 763 \\ \hline \Tstrut
*Decaps-640 1x & 4466 & 2152 & 1254 & 1 & 12.5 & 162 & 49 \\
*Decaps-640 4x & 4841 & 2661 & 1345 & 4 & 12.5 & 161 & 192 \\
*Decaps-640 8x & 5476 & 3479 & 1558 & 8 & 12.5 & 156 & 372 \\
*Decaps-640 16x & 6881 & 5081 & 1947 & 16 & 12.5 & 149 & 710 \\ \hline \Tstrut

Decaps-640 1x \cite{howe2018standard} & 7220 & 3549 & 1992 & 1 & 16 & 162 & 49 \\ \hline \Tstrut

Decaps-976 1x & 14217 & 2295 & 3956 & 1 & 0 & 188 & 25 \\
Decaps-976 4x & 16234 & 2853 & 4648 & 4 & 0 & 170 & 88 \\
Decaps-976 8x & 17451 & 3687 & 4985 & 8 & 0 & 161 & 167 \\
Decaps-976 16x & 18960 & 5285 & 5274 & 16 & 0 & 157 & 325 \\ \hline \Tstrut
*Decaps-976 1x & 4888 & 2153 & 1390 & 1 & 19 & 162 & 21 \\
*Decaps-976 4x & 5259 & 2662 & 1450 & 4 & 19 & 160 & 83 \\
*Decaps-976 8x & 5888 & 3490 & 1615 & 8 & 19 & 155 & 161 \\
*Decaps-976 16x & 7213 & 5087 & 2042 & 16 & 19 & 148 & 306 \\ \hline \Tstrut

Decaps-976 1x \cite{howe2018standard} & 7773 & 3559 & 2158 & 1 & 24 & 162 & 21 \\ \hline \Tstrut

\end{tabular}}
\end{table}

\begin{table}[!t]
\centering
\caption{FPGA resource consumption of the proposed FrodoKEM PRNG and Error Sampler designs. All results utilize a Xilinx Artix-7 FPGA.}\label{tab:otherresults}
%\vspace{-.2cm}
%\resizebox{0.7\textwidth}{!}{
\noindent\makebox[\textwidth]{
\begin{tabular}{l N{4}{0} N{4}{0} N{3}{0} N{1}{0} N{1}{0} N{3}{0} N{3}{0} }
\hline \Tstrut
  {\bf \textsc{FrodoKEM} Operation}
&  {\bf LUT} 
&  {\bf FF} 
&  {\bf Slices}  
&  {\bf DSP} 
&  {\bf BRAM}
&  {\bf MHz} 
&  {\bf Thr-Put}  \\ \hline \Tstrut
%\textbf{} & & & \textbf{BRAM} & &\textbf{Sec} \\ \hline

Error+Trivium & 401 & 311 & 179 & 0 & 0 & 211 & $211\text{m}$ \\
Trivium & 296 & 299 & 169 & 0 & 0 & 220 & $220\text{m}$ \\ \hline \Tstrut

Error+AES \cite{howe2018standard} & 1901 & 1140 & 756 & 0 & 0 & 184 & $184\text{m}$ \\
cSHAKE \cite{howe2018standard} & 2744 & 1685 & 766 & 0 & 0 & 172 & $1\text{m}$ \\ \hline \Tstrut

\end{tabular}}
\end{table}

\begin{figure}[h]
\centering
%    \advance\leftskip-1cm
\resizebox{0.8\textwidth}{!}{%
\input{figures/hw.tikz}
}
\caption{Visualisation of FPGA slice consumption of FrodoKEM's key generation, encapsulation, \& decapsulation on a Xilinx Artix-7. Decaps values overlap to show results with (*) and without BRAM.}
\label{fig:hw}
\end{figure}

@@

Since the majority of the proposed designs operate without BRAM, we were able to attain a much higher frequency than previous works. Overall our throughput outperforms previous comparable results, by factors between 1.13x and 1.19x~\cite{howe2018standard}. Moreover, whilst maintaining less area consumption than previous research we were able to increase the amount of parallel multipliers used. As a result, we can achieve up to 840 key generations per second (a 16.5x increase), 825 encapsulations per second (a 16.2x increase), and 710 operations per second (a 15.6x increase).

We also maintain the constant run-time which the previous implementation attains, as well as implementing first-order masking during decapsulation. The masking is also done using parallel multiplication and thus does not affect the run-time of the decapsulation module. The clock cycle counts for each module are easy to calculate; key generation requires $ (n^2\bar{n}) / k$ clocks, encapsulation requires $ (n^2\bar{n} + \bar{n}^2n)/k $ clocks, and decapsulation requires $(n^2\bar{n} + 2\bar{n}^2n)/k$ clocks, for dimensions $n=640$ or 976, $\bar{n}=8$, and $k$ referring to the number of parallel multipliers used.

\begin{figure}[h]
\centering
    \advance\leftskip-1cm
\resizebox{0.75\textwidth}{!}{%
\input{figures/hw_line.tikz}}
\caption{Comparison of the throughput performance per FPGA slice on a Xilinx Artix-7.}
 \label{fig:hw_line}
\end{figure}

>>>>

\vspace{-0.2cm}

\section{Results} \label{sec:results}
%One of the main contributions of this research is the proposal to use an alternative PRNG to the one used by Howe et al. \cite{howe2018standard}, cSHAKE. Comparatively, 
In this section we present the results obtained when implementing our FrodoKEM architecture. We provide a table of results for each of the key generation, encapsulation, and decapsulation designs in Tables \ref{tab:keyresults}, \ref{tab:encapsresults}, and \ref{tab:decapsresults}, respectively. We also provide results for the PRNG and Gaussian sampler in Table \ref{tab:otherresults}. All tables give comparative results of the previous FrodoKEM design in hardware, which utilize 1x LWE multiplier per clock cycle and completely conform to the FrodoKEM specifications by using cSHAKE where we are using Trivium. Moreover, all results are benchmarked on the same FPGA device as previous work, Xilinx Artix-7 XC7A35T FPGA, running on Vivado 2019.1.

\begin{table}[!t]
\centering
\caption{FPGA resource consumption of the proposed FrodoKEM \textbf{KeyGen} designs, using $k$ parallel multipliers, for both parameter sets, on a Xilinx Artix-7 FPGA.}\label{tab:keyresults}
%\vspace{-.2cm}
%\resizebox{0.7\textwidth}{!}{
%\noindent\makebox[\textwidth]{
\begin{tabular}{l r r r c c r}
\hline \Tstrut
  \textsc{FrodoKEM} & \multirow{2}{*}{LUT} & \multirow{2}{*}{FF} & \multirow{2}{*}{Slices} & DSP/ & \multirow{2}{*}{MHz} & Ops/ \\
Protocol &&&& BRAM && Sec  \\ \hline \Tstrut
%\textbf{} & & & \textbf{BRAM} & &\textbf{Sec} \\ \hline

KeyGen-640 1x & 971 & 433 & 290 & 1/0 & 191 & 59 \\
KeyGen-640 4x & 1174 & 781 & 355 & 4/0 & 185 & 226 \\
KeyGen-640 8x & 1679 & 1570 & 532 & 8/0 & 182 & 445 \\
KeyGen-640 16x & 2587 & 2994 & 855 & 16/0 & 172 & 840 \\ \hline \Tstrut
KeyGen-640 \cite{howe2018standard} & 3771 & 1800 & 1035 & 1/6 & 167 & 51 \\ \hline \Tstrut

KeyGen-976 1x & 1243 & 441 & 362 & 1/0 & 189 & 25 \\
KeyGen-976 4x & 1458 & 792 & 440 & 4/0 & 184 & 97 \\
KeyGen-976 8x & 1967 & 1576 & 617 & 8/0 & 178 & 187 \\
KeyGen-976 16x & 2869 & 3000 & 908 & 16/0 & 169 & 355 \\ \hline \Tstrut
KeyGen-976 \cite{howe2018standard} & 7139 & 1800 & 1939 & 1/8 & 167 & 22 \\ \hline \Tstrut

\end{tabular}%}
\end{table}

\begin{table}[!t]
\centering
\caption{FPGA resource consumption of the proposed FrodoKEM \textbf{Encapsulation} designs, using $k$ parallel multipliers, for both parameter sets, on a Xilinx Artix-7 FPGA.}\label{tab:encapsresults}
%\vspace{-.2cm}
%\resizebox{0.7\textwidth}{!}{
%\noindent\makebox[\textwidth]{
\begin{tabular}{l r r r c c r}
\hline \Tstrut
  \textsc{FrodoKEM} & \multirow{2}{*}{LUT} & \multirow{2}{*}{FF} & \multirow{2}{*}{Slices} & DSP/ & \multirow{2}{*}{MHz} & Ops/ \\
Protocol &&&& BRAM && Sec  \\ \hline \Tstrut
%\textbf{} & & & \textbf{BRAM} & &\textbf{Sec} \\ \hline

Encaps-640 1x & 4246 & 2131 & 1180 & 1/0 & 190 & 58 \\
Encaps-640 4x & 4620 & 2552 & 1338 & 4/0 & 183 & 221 \\
Encaps-640 8x & 5155 & 3356 & 1485 & 8/0 & 177 & 427 \\
Encaps-640 16x & 5796 & 4694 & 1692 & 16/0 & 171 & 825 \\ \hline \Tstrut
Encaps-640 \cite{howe2018standard} & 6745 & 3528 & 1855 & 1/11 & 167 & 51 \\ \hline \Tstrut

Encaps-976 1x & 4650 & 2118 & 1272 & 1/0 & 187 & 25 \\
Encaps-976 4x & 4996 & 2611 & 1455 & 4/0 & 180 & 94 \\
Encaps-976 8x & 5562 & 3349 & 1608 & 8/0 & 175 & 183 \\
Encaps-976 16x & 6188 & 4678 & 1782 & 16/0 & 168 & 350 \\ \hline \Tstrut

Encaps-976 \cite{howe2018standard} & 7209 & 3537 & 1985 & 1/16 & 167 & 22 \\ \hline \Tstrut

\end{tabular}%}
\end{table}

@@

There is a significant increase in area consumption of all the decapsulation results which do not utilize BRAM. This is mainly due to the need of storing public-key and secret-key matrices. We provide results for both architectures with and without BRAM. The design without BRAM has a significantly higher throughput, due to the much higher frequency. %These results are reported in Figure \ref{fig:hw_line}, which shows the efficiency of each design (namely their throughput) per FPGA slice utilized. 
Figure \ref{fig:hw} shows a slice count summary of all the proposed designs, showing a consistent and fairly linear increase in slice utilization as the number of parallel multipliers increases. We note on decapsulation results in Figure \ref{fig:hw} where the results would lie if BRAM is used, hence the total results for without BRAM include both red areas (i.e., they overlap). In most cases slice counts at least double for decapsulation when BRAM is removed, with only slight increases in throughout, hence it might be not be useful in some use cases. BRAM usage, however, is not as friendly when hardware designs are considered for ASIC, thus is it useful to consider designs both with and without BRAM.

\begin{table}[!t]
\centering
\caption{FPGA resource consumption of the proposed FrodoKEM \textbf{Decapsulation} designs, using $k$ parallel multipliers, for both parameter sets, on a Xilinx Artix-7 FPGA. *BRAM used}\label{tab:decapsresults}
%\vspace{-.2cm}
%\resizebox{0.7\textwidth}{!}{
%\noindent\makebox[\textwidth]{
\begin{tabular}{l r r r c c r}
\hline \Tstrut
  \textsc{FrodoKEM} & \multirow{2}{*}{LUT} & \multirow{2}{*}{FF} & \multirow{2}{*}{Slices} & DSP/ & \multirow{2}{*}{MHz} & Ops/ \\
Protocol &&&& BRAM && Sec  \\ \hline \Tstrut
%\textbf{} & & & \textbf{BRAM} & &\textbf{Sec} \\ \hline

Decaps-640 1x & 10518 & 2299 & 2933 & 1/0 & 190 & 57 \\
Decaps-640 4x & 11581 & 2818 & 3424 & 4/0 & 174 & 208 \\
Decaps-640 8x & 13128 & 3737 & 3710 & 8/0 & 164 & 391 \\
Decaps-640 16x & 14528 & 5335 & 4020 & 16/0 & 160 & 763 \\ \hline \Tstrut
*Decaps-640 1x & 4466 & 2152 & 1254 & 1/12.5 & 162 & 49 \\
*Decaps-640 4x & 4841 & 2661 & 1345 & 4/12.5 & 161 & 192 \\
*Decaps-640 8x & 5476 & 3479 & 1558 & 8/12.5 & 156 & 372 \\
*Decaps-640 16x & 6881 & 5081 & 1947 & 16/12.5 & 149 & 710 \\ \hline \Tstrut

Decaps-640 \cite{howe2018standard} & 7220 & 3549 & 1992 & 1/16 & 162 & 49 \\ \hline \Tstrut

Decaps-976 1x & 14217 & 2295 & 3956 & 1/0 & 188 & 25 \\
Decaps-976 4x & 16234 & 2853 & 4648 & 4/0 & 170 & 88 \\
Decaps-976 8x & 17451 & 3687 & 4985 & 8/0 & 161 & 167 \\
Decaps-976 16x & 18960 & 5285 & 5274 & 16/0 & 157 & 325 \\ \hline \Tstrut
*Decaps-976 1x & 4888 & 2153 & 1390 & 1/19 & 162 & 21 \\
*Decaps-976 4x & 5259 & 2662 & 1450 & 4/19 & 160 & 83 \\
*Decaps-976 8x & 5888 & 3490 & 1615 & 8/19 & 155 & 161 \\
*Decaps-976 16x & 7213 & 5087 & 2042 & 16/19 & 148 & 306 \\ \hline \Tstrut

Decaps-976 \cite{howe2018standard} & 7773 & 3559 & 2158 & 1/24 & 162 & 21 \\ \hline \Tstrut

\end{tabular}%}
\end{table}

\begin{table}[!t]
\centering
\caption{FPGA resource consumption of the proposed PRNG and Error Sampler designs on a Xilinx Artix-7 FPGA.}\label{tab:otherresults}
%\vspace{-.2cm}
%\resizebox{0.7\textwidth}{!}{
%\noindent\makebox[\textwidth]{
\begin{tabular}{l r r r c c r}
\hline \Tstrut
  \textsc{FrodoKEM} & \multirow{2}{*}{LUT} & \multirow{2}{*}{FF} & \multirow{2}{*}{Slices} & DSP/ & \multirow{2}{*}{MHz} & Ops/ \\
Protocol &&&& BRAM && Sec  \\ \hline \Tstrut
%\textbf{} & & & \textbf{BRAM} & &\textbf{Sec} \\ \hline

Error+Trivium & 401 & 311 & 179 & 0/0 & 211 & $211\text{m}$ \\
Trivium & 296 & 299 & 169 & 0/0 & 220 & $220\text{m}$ \\ \hline \Tstrut

Error+AES \cite{howe2018standard} & 1901 & 1140 & 756 & 0/0 & 184 & $184\text{m}$ \\
cSHAKE \cite{howe2018standard} & 2744 & 1685 & 766 & 0/0 & 172 & $1\text{m}$ \\ \hline \Tstrut

\end{tabular}%}
\end{table}

\begin{figure}[h]
\centering
%    \advance\leftskip-1cm
\resizebox{0.4\textwidth}{!}{%
\input{figures/hw.tikz}
}
\caption{Visualisation of FPGA slice consumption of FrodoKEM's key generation, encaps, and decaps on a Xilinx Artix-7. Decaps values overlap to show results with (*) and without BRAM.}
\label{fig:hw}
\end{figure}

@@

Since the majority of our proposed designs operate without BRAM, we are able to attain a higher frequency than previous works. Overall our throughput outperforms previous \emph{comparable} results, by factors between 1.13x and 1.19x~\cite{howe2018standard}. Moreover, whilst maintaining less area consumption than previous research, we are able to increase the amount of parallel multipliers. As a result, we can achieve up to 840 key generations per second (a 16.5x increase), 825 encapsulations per second (a 16.2x increase), and 710 operations per second (a 15.6x increase). We also maintain the constant run-time which the previous implementation attains, as well as implementing first-order masking during decapsulation. The masking is also done using parallel multiplication and thus does not affect the run-time of the decapsulation module. The clock cycle counts for each module are easy to calculate; key generation requires $ (n^2\bar{n}) / k$ clocks, encapsulation requires $ (n^2\bar{n} + \bar{n}^2n)/k $ clocks, and decapsulation requires $(n^2\bar{n} + 2\bar{n}^2n)/k$ clocks, for dimensions $n=640$ or 976, $\bar{n}=8$, and $k$ referring to the number of parallel multipliers used.

%\begin{figure}[h]
%\centering
%    \advance\leftskip-1cm
%\resizebox{0.45\textwidth}{!}{%
%\input{figures/hw_line.tikz}}
%\caption{Comparison of the throughput performance per FPGA slice on a Xilinx Artix-7.}
% \label{fig:hw_line}
%\end{figure}

### 2020-09-07 tex/abstract.tex
first commit

\begin{abstract} \label{astract}
FrodoKEM is a lattice-based key encapsulation mechanism, currently a semi-finalist in NIST's post-quantum standardization effort. A condition for these candidates is to use NIST standards for PRNG / seed-expanding, and as such most candidates utilize SHAKE, an XOF defined in the SHA-3 standard. However, for many of the candidates, this module is a significant implementation bottleneck. Trivium is a lightweight, ISO standard stream cipher which performs well in hardware and has been used in previous hardware designs for lattice-based cryptography. This research proposes optimized designs for FrodoKEM, concentrating on high throughput by parallelising the matrix multiplication operations within the cryptographic scheme. This process is eased by the use of Trivium due to its higher throughput and lower area consumption. The parallelisations proposed also complement the addition of first-order masking to the decapsulation module. Overall, we significantly increase the throughput of FrodoKEM; for encapsulation we see a 16x speed-up and for decapsulation we see a 14x speed-up compared to the previous state-of-the-art, whilst also maintaining a similar FPGA area footprint.

%achieving up to 840 key generations per second, 825 encapsulations per second, and 763 decapsulations per second, with almost no impact on the FPGA area consumption compared to the previous state-of-the-art.

%Recently, a standard lattice-based key encapsulation mechanism has been shown to have promising performance on FPGA. The cryptographic scheme, named FrodoKEM, is one of many candidates for NIST's call for post-quantum standardisation. This research proposes optimised designs for FrodoKEM, concentrating on high throughput by parallelising the matrix multiplication operations within the cryptographic scheme. This process is eased by the use of a much smaller and faster PRNG. The parallelisations proposed also complement the addition of first-order masking to the decapsulation module. Overall, we significantly increase the throughput of FrodoKEM, achieving up to 840 key generations per second, 825 encapsulations per second, and 763 decapsulations per second, with almost no impact on the FPGA area consumption compared to the previous state-of-the-art.

>>>>

\begin{abstract} \label{astract}
%FrodoKEM is a lattice-based key encapsulation mechanism, currently a semi-finalist in NIST’s post-quantum standardization effort. Most candidates utilize a NIST standard for randomness, a SHA-3 variant, however this is a significant implementation bottleneck. Trivium is a lightweight, ISO standard designed for hardware, having been used in previous implementation of lattice-based cryptography. This research proposes optimized designs for FrodoKEM, concentrating on high-throughput by parallelising its time-consuming matrix multiplication operations. The parallelisations proposed also complement the addition of first-order masking to decapsulation. We significantly increase \newline FrodoKEM’s throughput; with a $\sim$16x speed-up compared to the previous state-of-the-art, whilst also maintaining a similar FPGA area footprint.

FrodoKEM is a lattice-based key encapsulation mechanism, currently a semi-finalist in NIST’s post-quantum standardization effort. A condition for these candidates is to use NIST standards for PRNG /
seed-expanding, and as such most candidates utilize SHAKE, an XOF defined in the SHA-3 standard. However, for many of the candidates, this module is a significant implementation bottleneck. Trivium is a lightweight, ISO standard stream cipher which performs well in hardware and has been used in previous hardware designs for lattice-based cryptography. This research proposes optimized designs for FrodoKEM, concentrating on high throughput by parallelising the matrix multiplication operations within the cryptographic scheme. This process is eased by the use of Trivium due to its higher throughput and lower area consumption. The parallelisations proposed also complement the addition
of first-order masking to the decapsulation module. Overall, we significantly increase the throughput of FrodoKEM; for encapsulation we see a 16x speed-up, achieving 825 operations per second, and for decapsulation we see a 14x speed-up, achieving 763 operations per second, compared to the previous state-of-the-art, whilst also maintaining a similar FPGA area footprint of less than 2000 slices.
